---
title: "MTH211 Assignment 1 Part 1"
subtitle: "Analysing Academic Performance in Boards, JEE and IITK"
author: "Arqam Patel, Kumar Kanishk Singh, Aditya Kumar Tiwari"
format: pdf
editor: visual
toc: true
---

# Data collection

The data was collected using a [Google form](https://docs.google.com/forms/d/e/1FAIpQLSed7HoFO2fe1waPy8_l55DD_4hx3z9IlTguyRA7BvAb6YVOvA/viewform).

Based on the responses to these we tried to extract basic insights about possible correlation between academic performance in boards, JEE and IITK, as measured by percentage, JEE Advanced All India rank and CPI respectively.

It had the following questions:

1.  What was your percentage in class 10? (Continuous data: 0-100)
2.  What was your percentage in class 12? (Continuous data: 0-100)
3.  What is your current CPI? (Continuous data: 0-10)
4.  What was your JEE Mains rank (AIR) ? (Discrete, ranking data)
5.  What was your JEE Advanced rank (AIR)? (Discrete, ranking data)
6.  Which other exams did you appear for?
7.  Which Coaching did you take (if any)? (Nominal data)
8.  Which type of course did you attend at your coaching Institute? (Nominal data)
9.  Where did you take coaching? (Nominal data)
10. How many attempts did you take? (Ordinal data)
11. How many hours did you study on average daily during your JEE preparation? (Continuous data)
12. How many hours do you study on average daily currently? (Continuous data)
13. How would you compare your academic performance during JEE times vs In IIT? (Ordinal data)
14. How would you score yourself among fellow batchmates? (Discrete data)

The form received 34 responses, out of which 2 were discarded due to errors in filling the form like missing essential data.

```{r}
#| echo: false
data <- read.csv("MTH211A Form.csv")
summary(data)
```

# Form responses

## Class 10 percentage

We now look the mean, median, and the minimum and maximum values of class 10 percentage. We also look at the distribution's quartiles and check for outliers through a boxplot.

```{r}
#| echo: false
print("Average Class 10 Percentage:")
class10 <- as.numeric(data[,2])
print(mean(class10))
print("Median Class 10 Percentage:")
print(median(class10))
print("Minimum Class 10 Percentage:")
print(min(class10))
print("Maximum Class 10 Percentage:")
print(max(class10))

boxplot(class10, main = "Class 10 Percentage Boxplot", ylab = "Percentage")
```

The histogram does not taper at extremes, and is in fact quite flat, except a peak at 96-97%. This indicates that the distribution of 10th marks is quite different from those variables that do.

```{r}
#| echo: false

bins <- 88:100

hist(class10, bins, main = "Distribution of 10th %ages", cex.main = 1)

```

## Class 12 percentage

We now look the mean, median, and the minimum and maximum values of class 12 percentage. We also look at the distribution's quartiles and check for outliers through a boxplot.

```{r}
#| echo: false
print("Average Class 12 Percentage:")
class12 <- as.numeric(data[,3])
print(mean(class12))
print("Median Class 12 Percentage:")
print(median(class12))
print("Minimum Class 12 Percentage:")
print(min(class12))
print("Maximum Class 12 Percentage:")
print(max(class12))

boxplot(class12, main = "Current class 12 Percentage Boxplot", ylab = "Percentage")
```

We can also inspect the distribution through a histogram and see that it tapers at extremes. It is somewhat left skewed, with the modal class being 94-96.

```{r}
#| echo: false
hist(class12, main = "Distribution of 12th %ages", cex.main = 1)
```

## CPI 

We now look the mean, median, and the minimum and maximum values of CPI. We also look at the distribution's quartiles and check for outliers through a boxplot.

```{r}
#| echo: false
print("Average CPI:")
cpi <- as.numeric(data[,4])
print(mean(cpi))
print("Median CPI:")
print(median(cpi))
print("Minimum CPI:")
print(min(cpi))
print("Maximum CPI:")
print(max(cpi))

boxplot(cpi, main = "Current CPI Boxplot", ylab = "CPI")

```

We can also look at the distribution through a histogram and see that it is higher in the middle and tapers at extremes.

```{r}
#| echo: false

hist(cpi, main = "Distribution of CPIs", cex.main = 1)

```

## JEE Mains rank

Since this is discrete data that deals with ranking, we treat it differently than the last 3 fields.

```{r}
#| echo: false
main <- data[,5]

```

```{r}
#| echo: false



```

```{r}
#| echo: false



```

![](Screenshot%20(381).png)

![](Screenshot%20(382).png)

![](Screenshot%20(390).png)

![](Screenshot%20(383).png)

![](Screenshot%20(384).png)

![](Screenshot%20(385).png)

![](Screenshot%20(386).png)

```{r}
#| echo: false
prevHr <- as.numeric(data[-c(15,19,20,27,34),12])
print("The average study hours during JEE preparation were:")
print(mean(prevHr))
#The average study time has decreased to more than half
```

![](Screenshot%20(387).png)

```{r}
#| echo: false
currHr <- as.numeric(data[-c(11,15,19,20,34),13])
print("The current average study hours are:")
print(mean(currHr))
```

![](Screenshot%20(388).png)

Most of the students think that their performance has remained same or got worse.

![](Screenshot%20(389).png)

# Correlation analysis

## Class 12th and 10th percentages

```{r}
#| echo: false
print( paste("Pearson's correlation coefficient: "), cor(class12, class10, use="complete.obs"))
print( paste("Spearman's rank correlation: "), cor(class12, class10, use="complete.obs", method = "spearman"))

```

We observe that there is moderate correlation, according to both methods of measuring correlation, in 10th and 12th percentages.

## CPI and Class 10th percentage

```{r}
#| echo: false
print( paste("Pearson's correlation coefficient: "), cor(cpi, class10, use="complete.obs"))
print( paste("Spearman's rank correlation: "), cor(cpi, class10, use="complete.obs", method = "spearman"))
```

We again observe that there is moderate correlation, according to both methods of measuring correlation, between CPI and 10th percentages.

## CPI and Class 12th percentage

```{r}
#| echo: false
print( paste("Pearson's correlation coefficient: "), cor(cpi, class10, use="complete.obs"))
print( paste("Spearman's rank correlation: "), cor(cpi, class10, use="complete.obs", method = "spearman"))
```

We observe that there is negligible correlation, according to both methods of measuring correlation, between CPI and 10th percentages.

## JEE Main and Advance ranks

```{r}
#| echo: false
main <- as.numeric(data[,5])
adv <- as.numeric(data[,6])
print( paste("Spearman's rank correlation: "), cor(adv, main, use="complete.obs", method = "spearman"))
```

As expected, there is a relatively high degree of correlation between JEE Main and Advance ranks, measured according to Spearman's rank correlation.

## CPI and JEE Advanced ranks

```{r}
#| echo: false
print( paste("Spearman's rank correlation: "), cor(adv, cpi, use="complete.obs", method = "spearman"))
```

There is a low to moderate level correlation between CPI and JEE advance rank. The negative sign indicates that lower in magnitude JEE Advanced rank in general correlates to some extent with higher CPI.

## CPI and JEE Mains ranks

```{r}
#| echo: false
print( paste("Spearman's rank correlation: "), cor(main, cpi, use="complete.obs", method = "spearman"))

```

There is again, a low to moderate level correlation between CPI and JEE Mains rank. The negative sign indicates that lower in magnitude JEE Mains rank in general correlates to some extent with higher CPI. The correlation in this case is lower in magnitude than that of CPI with JEE Advance rank.
